---
title: "Assignement 10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Housekeeping
```{r}
library(modelr)
library(ggplot2)
library(tidyverse)
library(data.table)
library(purrr)
```

### Exercise 1: 
## Map each coefficient from mod1 and mod2 to a feature of the plot 
## with two facets. For instance, what is x1 in summaryd(mod2)? Where could you
## read it off (roughly) from the graph? Etc for x1:x2b and so on. If you get
## stuck, do ask for specific questions on Discourse. Correct answers for any 
## parameter look like this:
## x1 is the [slope/intercept/difference between slopes/intercepts of] for ... 
## Since it is [positive/negative] this means that ... is [larger/smaller] than ...

```{r}

mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)

predicted_data_mod1 <- sim3 %>% add_predictions(mod1) %>% add_residuals(mod1) %>% mutate(model = "mod1")
predicted_data_mod2 <- sim3 %>% add_predictions(mod2) %>% add_residuals(mod2) %>% mutate(model = "mod2")
predicted_data_combined <- rbind(predicted_data_mod1, predicted_data_mod2)

predicted_data_combined %>% ggplot(aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)

```

```{r}
mod1$coefficients
mod2$coefficients
```

**Ans:** Among the coefficient values from **mod1**; Intercept is the y-intercept of the line x2a and x1 is the slope of all of the 4 lines. 
x2b is the advantage/premium of each value in the line x2b over each value i x2a - when x2b is to added to intercept it equals the y-intercept of the line x2b (~4.75).
x2c is the advantage/premium of each value in the line x2c over each value i x2a - when x2c is to added to intercept it equals the y-intercept of the line x2c (~6.67).
x2d is the advantage/premium of each value in the line x2d over each value i x2a - when x2d is to added to intercept it equals the y-intercept of the line x2d (~4.23).

Among the coefficents from **mod2**; Intercept is the y-intercept of the line x2a and x1 is the slope of this line x2a. 
x2b is the advantage/premium of each value in the line x2b over each value i x2a - when x2b is to added to intercept it equals the y-intercept of the line x2b (~8.37). x1:x2b is the difference in the slopes of the lines x2a and x2b, when x1 is added to x1:x2b it forms the slope of the line x2b (~ -0.85)
x2c is the advantage/premium of each value in the line x2c over each value i x2a - when x2c is to added to intercept it equals the y-intercept of the line x2c (~5.73). x1:x2c is the difference in the slopes of the lines x2a and x2c, when x1 is added to x1:x2c it forms the slope of the line x2c (~ -0.03)
x2d is the advantage/premium of each value in the line x2d over each value i x2a - when x2d is to added to intercept it equals the y-intercept of the line x2d (~2.23). x1:x2d is the difference in the slopes of the lines x2a and x2d, when x1 is added to x1:x2d it forms the slope of the line x2d (~ 0.18)

### Exercise 2: 
## Do the faceting with gather_predictions and if needed with data_grid.
## Look at chapter 23 for help.

```{r}
grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  gather_predictions(mod1, mod2)

ggplot(sim3, aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)
```

### Exercise 3: Read/Skim 21.2, 21.3, and 21.4 so you are aware of some issues.
## Pick a short example from the notes that you feel you want to understand better
## and use some other use case to illustrate it (using the Vienna data, or 
## diamonds, or the same but in a different way.)

```{r}

files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)

read_files <- function(file){
  df <- read_csv(file)
}
read_files_df <- lapply(files, read_files)

merged_files_df <- rbindlist(read_files_df)

show_mean(merged_files_df)
show_mean <- function(df) {
output <- vector("double", ncol(df))
          
for(i in seq_along(df)){
  if(is_numeric(df[[i]])){
    output[[i]] <- mean(df[[i]])
  }
}
print("((((((((((((((((((((((((((((((")
for(i in seq_along(output)){
  if(!is.na(output[[i]])){
  cat(paste0(names(df)[i], ": ", round(mean(df[[i]]), 2)), fill = TRUE)
    }
  }
}

```
