---
title: "Data Tasks"
output: pdf_document
---

```{r setup, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

### Chapter 2 - Data Exercise Q.5

Table showing the number of seasons each team spent in the English Premier League

```{r, echo=FALSE, results='hide'}

epl_data <- read_csv('epl_games.csv', col_names = TRUE) 
seasons_played_in_epl <-  epl_data %>%
    group_by(team_home) %>%
    mutate(no_of_seasons = n_distinct(season)) %>%
distinct(team_home, no_of_seasons) %>%
rename(Team = team_home, No_Of_Seasons_In_Epl = no_of_seasons)
seasons_played_in_epl

```

Teams that played all 11 seasons in the EPL

```{r}
seasons_played_in_epl %>% filter(No_Of_Seasons_In_Epl == 11) %>% select(Team)
```
Teams that played only once in the EPL

```{r}
seasons_played_in_epl %>% filter(No_Of_Seasons_In_Epl == 1) %>% select(Team)

```


### Chapter 3 - Data Exercise Question 3


```{r}
library(scales)
home_adv_2018 <- epl_data %>% filter(season == 2018) 


  home_adv_2018 %>% ggplot(mapping = aes(x = goals_home - goals_away)) + geom_bar(aes(y = (..count..)/sum(..count..)), color = "Dark Blue", fill = "Sky Blue") +
  geom_text(aes(y = ((..count..)/sum(..count..)), label = scales::percent((..count..)/sum(..count..))), stat = "count", vjust = -0.25) +
  scale_y_continuous(labels = percent) + labs(title = "Home Advantage in 2018/19", y = "Percentage of Games", x = "Goal Difference")

home_adv_2018 %>% summarise(No_of_observations = n(), 
                              Mean = round(mean(goals_home - goals_away), 2), 
                              standard_dev = round(sd(goals_home - goals_away), 2), 
                              Percent_positive = round(sum(goals_home - goals_away > 0)/n(), 2)*100, 
                              Percent_zero = round(sum(goals_home - goals_away == 0)/n(), 2)*100, 
                              Percent_negative = round(sum(goals_home - goals_away < 0)/n(), 2)*100)
```

**Conclusion:** We plotted a frequency distribution of Goal Difference in the 2018/19 season and found out that the mode was higher in 2018 compared to 2017 (in the book). The mean and standard deviation are also slightly higher in 2018, while the percentage of matches with negative goal difference is also higher in 2018 and the percentage of matches with zero goal difference is lower in 2018 compared to 2016.

### Chapter 4 - Data Exercise Question 2.

```{r}
home_totals <- epl_data %>% 
  filter(season == 2017) %>% 
  group_by(team_home) %>% 
  summarise(total_points = sum(points_home)) %>% 
  arrange(total_points) %>% 
  rename(Team = team_home)
  
away_totals <- epl_data %>% 
  filter(season == 2017) %>% 
  group_by(team_away) %>% 
  summarise(total_points = sum(points_away)) %>% 
  arrange(total_points) %>% 
  rename(Team = team_away)

season_totals <- merge(home_totals, away_totals, by = "Team")
season_totals <- season_totals %>% 
  mutate(total_points = total_points.x + total_points.y) %>% 
  arrange(desc(total_points))

season_2017_binned <- mutate(season_totals, 
                             bin = cut(season_totals$total_points, c(-Inf, 54, 42, Inf), 
                                       labels = c("Relegation Battle", "Mid-Table", "Top Six")))
season_2017_binned <- select(season_2017_binned, Team, bin)
season_2018 <- epl_data %>% 
  filter(season == 2018) %>% rename(Team = team_home)
season_2018_binned <- merge(season_2017_binned, season_2018, 
                            by = "Team", all.y = TRUE)
season_2018_binned$bin <- replace(season_2018_binned$bin, 
                                  is.na(season_2018_binned$bin), 
                                  "Relegation Battle")

```

2018 Statistics with Team sorted into three bins; Top Six, Mid-Table and Relegation Battle according to their in the previous season

```{r}
season_2018_binned %>% distinct(Team, bin) %>% arrange(bin)

season_2018_binned %>%  
  group_by(bin) %>% summarise(No_of_observations = n(), 
                              Mean = round(mean(goals_home - goals_away), 2), 
                              standard_dev = round(sd(goals_home - goals_away), 2), 
                              Percent_positive = round(sum(goals_home - goals_away > 0)/n(), 2)*100, 
                              Percent_zero = round(sum(goals_home - goals_away == 0)/n(), 2)*100, 
                              Percent_negative = round(sum(goals_home - goals_away < 0)/n(), 2)*100)
```

Home Advantage in our three bins


1. Histogram 

```{r}
topSix <- season_2018_binned %>% filter(bin == "Top Six") 

topSix_adv <- topSix %>% 
  ggplot(mapping = aes(x = goals_home - goals_away)) + 
  geom_histogram() + 
  labs(title = "Home advantage amongst Top Six",
                          x = "Goal Difference")

midTable <- season_2018_binned %>% filter(bin == "Mid-Table")


midTable_adv <- midTable %>%
  ggplot(mapping = aes(x = goals_home - goals_away)) + 
  geom_histogram() + 
  labs(title = "Home advantage amongst Mid-Table",
                          x = "Goal Difference")

releBattle <- season_2018_binned %>% filter(bin == "Relegation Battle") 

releBattle_adv <- releBattle %>% 
  ggplot(mapping = aes(x = goals_home - goals_away))  + 
  geom_histogram() + 
  labs(title = "Home advantage amongst the teams in Relegation Battle",
                          x = "Goal Difference")

ggarrange(topSix_adv, midTable_adv, releBattle_adv, vjust = -3)
```

2. Bin Scatter

```{r}
x <- season_2018_binned %>%  
  group_by(bin) %>% 
  summarise(goal_diff = mean(goals_home - goals_away), 
            count_of_teams = n_distinct(Team), goals_home = mean(goals_home), 
            goals_away = mean(goals_away)) 

x %>% ggplot(mapping = aes(x = goals_home , y = goals_away, color = bin, size = goal_diff)) + 
  geom_point() + labs(title = "Home advantage amongst different bins")

```

3. Scatter Plot

```{r}
y <- season_2018_binned %>% 
  group_by(bin, Team) %>% 
  summarise(goal_diff = mean(goals_home - goals_away), 
            goals_home = mean(goals_home), goals_away = mean(goals_away))
y %>% ggplot(mapping = aes(x = goals_home , y = goals_away, 
                           color = bin, size = goal_diff)) + 
  geom_point(labels = season_2018_binned$Team) +
  labs(title = "Scatter Plot show-casing home advantage among different bins")
```

**Conclusion:** As can be seen by the two scatter plots, home advantage is most significant for the Top Six Teams, where they score around 2 home goals for each goal the away team scores. This advantage is very mild for mid_table teams where they score 1.32 home goals for every 1.28 away goals. Finally this pattern/advantage is inversed among teams that battle for relegation - these teams score 1 goal for about 1.5 every goal they concede at home.





